Comments on Latent Ranking

### 2021-05-21 output_scripts/revisions_may.Rmd
styler rmd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      # fig.path = "paper_figure/",
                      # dev = 'png',
                      # fig.height = 4.5,
                      # fig.width = 7,
                      warning = FALSE)

@@

library(here)
library(rstan)
library(ppdiag)
Sys.setenv(LOCAL_CPPFLAGS = '-march=corei7 -mtune=corei7')
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(R.utils)
library(RColorBrewer)
library(fields)
library(bayesplot)
library(viridis)
library(here)
library(compete)
library(colorspace)
source(paste(func_dir,'naiveRankHierarchy.R',sep = ""))
source(paste(func_dir,'expertRankHierarchy.R',sep = ""))
source(paste(func_dir,'cleanData.R',sep = "")) 
source(paste(func_dir,'prepareDataStan.R',sep = "")) 
source(paste(func_dir,'inferLatentMmhp.R',sep = ""))
source(paste(func_dir,'plotUtil.R',sep = "")) 
source(paste(func_dir,'mmhp.R',sep = ""))
source(paste(func_dir,'uniHawkes.R',sep = ""))
source(paste(func_dir,'simulatePrediction.R',sep = ""))
source(paste(func_dir,'myGlicko.R',sep = ""))
source(paste(func_dir,'matrixPlotParameter.R',sep = ""))
source(paste(func_dir,'residualStructureScore.R',sep = ""))
source(paste(func_dir,'drawIntensity.R',sep =""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")
```

@@

model1_fn <- list(alpha.fun = function(x, y, eta1, eta2, eta3){
  return(eta1 * x * y * exp(-eta2 * abs(x-y))/(1 + exp(-eta3 *(x-y))))})

model3_fn <- list(alpha.fun = function(x, y, eta1, eta2){
  return(eta1*x*y*exp(-eta2*abs(x-y)))},
  q1.fun = function(x,y,eta3){return(exp(-eta3*x))},
  q0.fun = function(x,y,eta3){return(exp(-eta3*y))})

@@

object_fn <- list(alpha.fun = function(x,y,eta1,eta2){
  return(eta1*x*y*exp(-eta2*abs(x-y)))},
  q1.fun = function(x,y,eta3){return(exp(-eta3*x))},
  q0.fun = function(x,y,eta3){return(exp(-eta3*y))})

object_par <- list(sim_lambda_1 = 0.2,
                   sim_eta_1 = 1.5, # this has to be < beta
                   gamma_var = runif(n = num_nodes, min = 0.01, max = 0.05),
                   zeta_var = runif(n = num_nodes, min = 0.01, max = 0.05),
                   sim_eta_2 = 0.6,
                   sim_eta_3 = 3,
                   sim_beta = 2,
                   f_vec_1 = seq(from = 0.05, to = 0.95,
                                 length.out = num_nodes))

object_matrix <- list(lambda0_matrix=outer(object_par$gamma_var,
                                           object_par$zeta_var, "+"),
                      lambda1_matrix=matrix(object_par$sim_lambda_1,
                                            nrow=length(object_par$f_vec_1),
                                            ncol=length(object_par$f_vec_1)),
                      alpha_matrix=formMatrix(function(x,y)
                        object_fn$alpha.fun(x,y,object_par$sim_eta_1,
                                            object_par$sim_eta_2),
                        object_par$f_vec_1),
                      beta_matrix=matrix(object_par$sim_beta,
                                         nrow=length(object_par$f_vec_1),
                                         ncol=length(object_par$f_vec_1)),
                      q1_matrix=formMatrix(function(x,y)
                        object_fn$q1.fun(x, y,
                                         object_par$sim_eta_3),
                        object_par$f_vec_1),
                      q2_matrix=formMatrix(function(x,y)
                        object_fn$q0.fun(x, y,
                                         object_par$sim_eta_3),
                        object_par$f_vec_1))

## Simulate
sim_model3_data <- list()
N_array <- array(0, c(1, num_nodes, num_nodes))
# for(i in c(1:n_sim)){
sim_model3_data <- simulateLatentMMHP(lambda0_matrix =
                                        object_matrix$lambda0_matrix,
                                      lambda1_matrix =
                                        object_matrix$lambda1_matrix,
                                      alpha_matrix = object_matrix$alpha_matrix,
                                      beta_matrix = object_matrix$beta_matrix,
                                      q1_matrix = object_matrix$q1_matrix,
                                      q2_matrix = object_matrix$q2_matrix,
                                      horizon = obs_time)
clean_sim_data <- cleanSimulationData(raw_data = sim_model3_data,
                                      cut_off = cut_off,
                                      N = length(object_par$f_vec_1))
N_array <- clean_sim_data$N_count

```

@@

obj <- pp_mmhp(lambda0 = object_matrix$lambda0_matrix[start, end],
        lambda1 = object_matrix$lambda1_matrix[start, end],
        alpha = object_matrix$alpha_matrix[start, end],
        beta = object_matrix$beta_matrix[start, end],
        Q = matrix(c(-q1, q1, q2, -q2), nrow = 2, ncol = 2, byrow = TRUE))

@@

```

```{r ranking_sim_data}
### i&si method
count_data_dc <- get_wl_matrix(df = cbind(clean_sim_data$start,
                                          clean_sim_data$end))
isi_dc.out <- compete::isi98(m = count_data_dc, random = TRUE)
isi_dc.out$best_order

### agg rank method
agg_rank_data <- clean_sim_data$N_count
agg_rank_model <- stan_model(here("lib","latent_rank_agg_sim.stan"))

agg_rank_fit <- rstan::sampling(agg_rank_model,
                                data = list(n = num_nodes,
                                            n_matrix = agg_rank_data),
                                iter = 1000,
                                chains = 4)
agg_sims <- rstan::extract(agg_rank_fit)

agg_rank <- order(apply(agg_sims$x, 2, mean))
agg_rank

### then glicko
glicko_data <- tibble(start = clean_sim_data$start,
                      end = clean_sim_data$end)

@@

library(PlayerRatings)
gl_train <- my_glicko(glicko_data, history=TRUE, cval=2)

@@

gl_ranks <- order(gl_train$ratings$Rating)
gl_ranks
```

@@

myMultiMatrixPlot(X = matrix_lst_plot,
                  no_matrix = 2,
                  n_row = 1,
                  xLabels = 1:20,
                  yLabels = 20:1,
                  min = -30,
                  max = 30,
                  axis_cex = 2,
                  title_cex = 1.8,
                  colorPalette = "RdBu",
                  if.emp = FALSE,
                  #legend.mar=c(0.5,.5,0.5,0.5),
                  title_lst = list("Raw","Pearson"), 
                  #title_lst = list("I-MMHP"),
                  col_axis = c(-30,-15,0,15,30), 
                  fake_matrix = TRUE,
                  matrix.mar = c(2.5,2.5,2.5,1) )

```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  # fig.path = "paper_figure/",
  # dev = 'png',
  # fig.height = 4.5,
  # fig.width = 7,
  warning = FALSE
)

@@

library(here)
library(rstan)
library(ppdiag)
Sys.setenv(LOCAL_CPPFLAGS = "-march=corei7 -mtune=corei7")
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(R.utils)
library(RColorBrewer)
library(fields)
library(bayesplot)
library(viridis)
library(here)
library(compete)
library(colorspace)
source(paste(func_dir, "naiveRankHierarchy.R", sep = ""))
source(paste(func_dir, "expertRankHierarchy.R", sep = ""))
source(paste(func_dir, "cleanData.R", sep = ""))
source(paste(func_dir, "prepareDataStan.R", sep = ""))
source(paste(func_dir, "inferLatentMmhp.R", sep = ""))
source(paste(func_dir, "plotUtil.R", sep = ""))
source(paste(func_dir, "mmhp.R", sep = ""))
source(paste(func_dir, "uniHawkes.R", sep = ""))
source(paste(func_dir, "simulatePrediction.R", sep = ""))
source(paste(func_dir, "myGlicko.R", sep = ""))
source(paste(func_dir, "matrixPlotParameter.R", sep = ""))
source(paste(func_dir, "residualStructureScore.R", sep = ""))
source(paste(func_dir, "drawIntensity.R", sep = ""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")
```

@@

model1_fn <- list(alpha.fun = function(x, y, eta1, eta2, eta3) {
  return(eta1 * x * y * exp(-eta2 * abs(x - y)) / (1 + exp(-eta3 * (x - y))))
})

model3_fn <- list(
  alpha.fun = function(x, y, eta1, eta2) {
    return(eta1 * x * y * exp(-eta2 * abs(x - y)))
  },
  q1.fun = function(x, y, eta3) {
    return(exp(-eta3 * x))
  },
  q0.fun = function(x, y, eta3) {
    return(exp(-eta3 * y))
  }
)

@@

object_fn <- list(
  alpha.fun = function(x, y, eta1, eta2) {
    return(eta1 * x * y * exp(-eta2 * abs(x - y)))
  },
  q1.fun = function(x, y, eta3) {
    return(exp(-eta3 * x))
  },
  q0.fun = function(x, y, eta3) {
    return(exp(-eta3 * y))
  }
)

object_par <- list(
  sim_lambda_1 = 0.2,
  sim_eta_1 = 1.5, # this has to be < beta
  gamma_var = runif(n = num_nodes, min = 0.01, max = 0.05),
  zeta_var = runif(n = num_nodes, min = 0.01, max = 0.05),
  sim_eta_2 = 0.6,
  sim_eta_3 = 3,
  sim_beta = 2,
  f_vec_1 = seq(
    from = 0.05, to = 0.95,
    length.out = num_nodes
  )
)

object_matrix <- list(
  lambda0_matrix = outer(
    object_par$gamma_var,
    object_par$zeta_var, "+"
  ),
  lambda1_matrix = matrix(object_par$sim_lambda_1,
    nrow = length(object_par$f_vec_1),
    ncol = length(object_par$f_vec_1)
  ),
  alpha_matrix = formMatrix(
    function(x, y) {
      object_fn$alpha.fun(
        x, y, object_par$sim_eta_1,
        object_par$sim_eta_2
      )
    },
    object_par$f_vec_1
  ),
  beta_matrix = matrix(object_par$sim_beta,
    nrow = length(object_par$f_vec_1),
    ncol = length(object_par$f_vec_1)
  ),
  q1_matrix = formMatrix(
    function(x, y) {
      object_fn$q1.fun(
        x, y,
        object_par$sim_eta_3
      )
    },
    object_par$f_vec_1
  ),
  q2_matrix = formMatrix(
    function(x, y) {
      object_fn$q0.fun(
        x, y,
        object_par$sim_eta_3
      )
    },
    object_par$f_vec_1
  )
)

## Simulate
sim_model3_data <- list()
N_array <- array(0, c(1, num_nodes, num_nodes))
# for(i in c(1:n_sim)){
sim_model3_data <- simulateLatentMMHP(
  lambda0_matrix =
    object_matrix$lambda0_matrix,
  lambda1_matrix =
    object_matrix$lambda1_matrix,
  alpha_matrix = object_matrix$alpha_matrix,
  beta_matrix = object_matrix$beta_matrix,
  q1_matrix = object_matrix$q1_matrix,
  q2_matrix = object_matrix$q2_matrix,
  horizon = obs_time
)
clean_sim_data <- cleanSimulationData(
  raw_data = sim_model3_data,
  cut_off = cut_off,
  N = length(object_par$f_vec_1)
)
N_array <- clean_sim_data$N_count
```

@@

obj <- pp_mmhp(
  lambda0 = object_matrix$lambda0_matrix[start, end],
  lambda1 = object_matrix$lambda1_matrix[start, end],
  alpha = object_matrix$alpha_matrix[start, end],
  beta = object_matrix$beta_matrix[start, end],
  Q = matrix(c(-q1, q1, q2, -q2), nrow = 2, ncol = 2, byrow = TRUE)
)

@@

obj$q1 <- q1
obj$q2 <- q2

viterbi_result <- myViterbi(
  events = events,
  param = obj,
  termination = obs_time
)

latent_inter <- interpolateLatentTrajectory(obj,
  events,
  viterbi_result$zt_v,
  initial.state =
    viterbi_result$initial_state,
  termination.time = obs_time,
  termination.state =
    viterbi_result$termination_state
)

event_state <- ppdiag:::mmhp_event_state(params = obj, events = events)

latent_state <- ppdiag:::interpolate_mmhp_latent(
  params = obj, events,
  zt = event_state$zt
)

mmhp_latent <- list()
mmhp_latent$start <- 0
mmhp_latent$end <- obs_time
mmhp_latent$x <- c(0, latent_state$x.hat, obs_time)
mmhp_latent$z <- c(viterbi_result$initial_state, latent_state$z.hat)
mmhp_latent$events <- c(0, events)
mmhp_latent$zt <- c(1, event_state$zt)
ppdiag::drawUniMMHPIntensity(obj, mmhp_latent)

# this plot is not correct for later Hawkes states...
```

```{r ranking_sim_data}
### i&si method
count_data_dc <- get_wl_matrix(df = cbind(
  clean_sim_data$start,
  clean_sim_data$end
))
isi_dc.out <- compete::isi98(m = count_data_dc, random = TRUE)
as.numeric(isi_dc.out$best_order) # from highest to lowest

### agg rank method
agg_rank_data <- clean_sim_data$N_count
agg_rank_model <- stan_model(here("lib", "latent_rank_agg_sim.stan"))

agg_rank_fit <- rstan::sampling(agg_rank_model,
  data = list(
    n = num_nodes,
    n_matrix = agg_rank_data
  ),
  iter = 1000,
  chains = 4
)
agg_sims <- rstan::extract(agg_rank_fit)

agg_rank <- order(apply(agg_sims$x, 2, mean))
rev(agg_rank) # from highest to lowest

### then glicko
glicko_data <- tibble(
  start = clean_sim_data$start,
  end = clean_sim_data$end
)

@@

library(PlayerRatings)
gl_train <- my_glicko(glicko_data, history = TRUE, cval = 2)

@@

gl_ranks <- order(gl_train$ratings$Rating)
gl_ranks # highest to lowest
```

@@

myMultiMatrixPlot(
  X = matrix_lst_plot,
  no_matrix = 2,
  n_row = 1,
  xLabels = 1:20,
  yLabels = 20:1,
  min = -30,
  max = 30,
  axis_cex = 2,
  title_cex = 1.8,
  colorPalette = "RdBu",
  if.emp = FALSE,
  # legend.mar=c(0.5,.5,0.5,0.5),
  title_lst = list("Raw", "Pearson"),
  # title_lst = list("I-MMHP"),
  col_axis = c(-30, -15, 0, 15, 30),
  fake_matrix = TRUE,
  matrix.mar = c(2.5, 2.5, 2.5, 1)
)
```

## Explaining the g function

```{r plot_g}
alpha.fun <- function(x, y, eta1, eta2) {
  return(eta1 * x * y * exp(-eta2 * abs(x - y)))
}

f_vec <- c(0.1, 0.2, 0.4, 0.8, 0.9)

eta1 <- 6.6
eta2 <- 0.37

crossing(send = f_vec, rec = f_vec) %>%
  filter(send != rec) %>%
  mutate(alpha = alpha.fun(send, rec, eta1, eta2))
```

## Community Structure Plot

```{r recreate_jing_figure}
load("../output/state_separation_plot.RData")
for (current_cohort in c(3)) {
  total_event_array <- total_event_array_lst[[current_cohort]]
  active_event_array <- active_event_array_lst[[current_cohort]]
  clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
  return_df <- cleanObservationPeriod(current_cohort, clean_data)

  start_win <- floor(max(return_df$observe.id) / 3) * (0:2) + 1
  end_win <- floor(max(return_df$observe.id) / 3) * (1:3)
  matrix1 <- apply(
    total_event_array[, , start_win[1]:end_win[1]] -
      active_event_array[, , start_win[1]:end_win[1]],
    c(1, 2), sum
  )
  matrix2 <- apply(
    total_event_array[, , start_win[2]:end_win[2]] -
      active_event_array[, , start_win[2]:end_win[2]],
    c(1, 2), sum
  )
  matrix3 <- apply(
    total_event_array[, , start_win[3]:end_win[3]] -
      active_event_array[, , start_win[3]:end_win[3]],
    c(1, 2), sum
  )

  g_community1 <- graph_from_adjacency_matrix(matrix1, mode = "undirected")
  lec1 <- cluster_leading_eigen(g_community1)

  g_community2 <- graph_from_adjacency_matrix(matrix2, mode = "undirected")
  lec2 <- cluster_leading_eigen(g_community2)

  g_community3 <- graph_from_adjacency_matrix(matrix3, mode = "undirected")
  lec3 <- cluster_leading_eigen(g_community3)

  ### Make plot for t=1
  png(paste(plot_path, "real_clustering_1.png", sep = ""), 
      height = 600,
      width = 600)
  myCircularPlot(
    cur.matrix = matrix1,
    cur.order = c(1, 4, 10, 5, 11, 6, 7, 2, 8, 12, 9, 3),
    cur.color = c(
      colorRampPalette(brewer.pal(
        name = "YlOrBr",
        n =
          8
      ))(14)[c(
        13,
        9,
        5
      )],
      colorRampPalette(brewer.pal(
        name = "PuRd",
        n =
          8
      ))(15)[c(
        4,
        6,
        8,
        10,
        12,
        14
      )],
      colorRampPalette(brewer.pal(
        name = "Blues",
        n =
          8
      ))(10)[c(
        5,
        7,
        9
      )]
    ),
    cur.gap = c(rep(5, 2), 30, rep(5, 5), 30, rep(5, 2), 30)
  )
  dev.off()

  ### Make plot for t=2
  png(paste(plot_path, "real_clustering_2.png", sep = ""),
   height = 600,
   width = 600
 )
 myCircularPlot(
   cur.matrix = matrix2,
   cur.order = c(1, 7, 8, 2, 3, 4, 5, 9, 10, 6, 11, 12),
   cur.color = c(
     colorRampPalette(brewer.pal(
       name = "YlOrBr",
       n = 8))(14)[c(13, 11, 9, 7, 5, 3)],
     colorRampPalette(brewer.pal(
       name = "Blues",
       n = 8))(15)[c(4, 6, 8, 10, 12, 14)]
   ),
   cur.gap = c(rep(5, 5), 30, rep(5, 5), 30)
 )
 dev.off()

  ### Make plot for t=3
  png(paste(plot_path, "real_clustering_3.png", sep = ""),
      height = 600,
      width = 600)
  myCircularPlot(
    cur.matrix = matrix3[-3, -3],
    cur.order = c(1, 10, 5, 2, 3, 6, 7, 8, 4, 11, 9),
    cur.color = c(
      colorRampPalette(
        brewer.pal(name = "YlOrBr",
                   n = 8))(14)[c(13, 9, 5, 3)],
      colorRampPalette(brewer.pal(
        name = "Blues",
        n = 8))(16)[c(7, 9, 11, 13, 15)],
      colorRampPalette(brewer.pal(
        name = "PuRd",
        n = 8))(15)[c(8, 12)]
    ),
    cur.gap = c(rep(5, 3), 30, rep(5, 4), 30, rep(5, 1), 30),
    cur.ID = c(1:12)[-3]
  )
  dev.off()
}
```
